High-level cognition is supported by information-rich but compressible brain activity patterns

### 2023-12-24 paper/main.tex
added network permutation analyses + t-tests, report results in paper

If the informativeness and compressibility of brain activity patterns vary over
time, do these properties might also vary across brain networks? We used a
network parcellation identified by \cite{YeoEtal11} to segment the brain into
seven distinct networks. The networks can be sorted (roughly) in order from
lower-level to higher-level cortex as follows (Figs.~\ref{fig:networks}A--C):
visual, somatomotor, dorsal attention, ventral attention, limbic,
frontoparietal, and default mode. Next, we computed decoding curves separately
for the activity patterns within each network and identified each network's
inflection point, for each experimental condition. Moving from low-order
networks to higher-order networks, we found that decoding accuracy tended to
increase in the higher-level experimental conditions and decrease (slightly) in
the lower-level experimental conditions (Fig.~\ref{fig:networks}D, E;
Spearman's rank correlation between decoding accuracy and network order:
intact: $\rho = 0.362, p < 0.001$; paragraph: $\rho = 0.441, p < 0.001$; word:
$\rho = -0.102, p = 0.007$; rest: $\rho = -0.354, p < 0.001$). This suggests
that higher-order networks may carry more content-relevant or stimulus-driven
``information.'' We found no clear trends in the proportions of components
required to achieve 5\% decoding accuracy across networks or conditions
(Fig.~\ref{fig:networks}F). We note that the limbic network we considered here
often overlaps with low (imaging) signal regions, and therefore it may be
difficult to draw strong conclusions about this network's informativeness or
compressibility. We also considered the possibility that the correlations with
network order might be influenced by the numbers of nodes in each network. We
designed a permutation-based procedure to address this possibility, whereby we
repeated the above analyses using shuffled network labels (see \textit{Network
permutation tests}).  The correlations between decoding accuracy and network
order... <INSERT RESULTS>

@@

>>>>

If the informativeness and compressibility of brain activity patterns vary over
time, do these properties might also vary across brain networks? We used a
network parcellation identified by \cite{YeoEtal11} to segment the brain into
seven distinct networks. The networks can be sorted (roughly) in order from
lower-level to higher-level cortex as follows (Figs.~\ref{fig:networks}A--C):
visual, somatomotor, dorsal attention, ventral attention, limbic,
frontoparietal, and default mode. Next, we computed decoding curves separately
for the activity patterns within each network and identified each network's
inflection point, for each experimental condition. Moving from low-order
networks to higher-order networks, we found that decoding accuracy tended to
increase in the higher-level experimental conditions and decrease (slightly) in
the lower-level experimental conditions (Fig.~\ref{fig:networks}D, E;
Spearman's rank correlation between decoding accuracy and network order:
intact: $\rho = 0.362, p < 0.001$; paragraph: $\rho = 0.441, p < 0.001$; word:
$\rho = -0.102, p = 0.007$; rest: $\rho = -0.354, p < 0.001$). This suggests
that higher-order networks may carry more content-relevant or stimulus-driven
``information.'' We found no clear trends in the proportions of components
required to achieve 5\% decoding accuracy across networks or conditions
(Fig.~\ref{fig:networks}F). We note that the limbic network we considered here
often overlaps with low (imaging) signal regions, and therefore it may be
difficult to draw strong conclusions about this network's informativeness or
compressibility. We also considered the possibility that the correlations with
network order might be influenced by the numbers of nodes in each network. We
designed a permutation-based procedure to address this possibility, whereby we
repeated the above analyses using shuffled network labels (see \textit{Network
permutation tests}). The correlations between decoding accuracy and network
order were reliably more positive than the shuffled correlations for the intact
($t(1998) = 276.431, p < 0.001$) and paragraph ($t(1998) = 330.334, p < 0.001$)
conditions, and reliably more negative for the word ($t(1998) = -16.386, p <
0.001$) and rest ($t(1998) = -318.631, p < 0.001$) conditions. These results
suggest that the correlations between decoding accuracy and network order were
not driven solely by the numbers of nodes in each network.

@@

\subsection*{Network permutation tests}

In our analyses of how informativeness varied across brain networks, we
considered the possibility that the correlations with network order might be
influenced by the numbers of nodes in each network. We designed a
permutation-based procedure to address this possibility, whereby we repeated
the above analyses using shuffled network labels. Specifically, for each of $n_1 =
10$ iterations, we randomly shuffled (without replacement) the network labels of
the HTFA nodes, and then we re-ran our entire decoding analysis pipeline,
including applying PCA with $3...m$ features for each condition (where $m$ is the
number of nodes in the given network), and then running 100 cross-validation
runs of the decoding procedure for each condition and number of components.
This resulted in 10 sets of shuffled data where each network had the same
numbers of nodes, but where the decoding results no longer maintained the
fidelity of each individual network.

We sampled the original and shuffled datasets (with replacement) to create $n_2
= 1000$ bootstrap samples. For each bootstrap sample, we computed the
correlations between the decoding accuracies and network order for each
condition and number of components. This yielded a distribution of $n_2$
correlation values for each condition, for both the original and shuffled
datasets. We then compared the distributions of Spearman's $\rho$ values for
the original and shuffled datasets using a two-sided independent samples
Welch's $t$-tests.
